Self Regulation Ontology Retest Data Report

### 2017-04-23 SRO_Report.Rmd
mostly cleaned up refactoring calculation of reliability df

```{r}
cor_df <- data.frame(spearman = rep(NA, length(numeric_cols)))

row.names(cor_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
  cor_df[numeric_cols[i], 'spearman'] <- get_spearman(numeric_cols[i]) 
}

#Calculate ICC df
icc_df <- data.frame(icc = rep(NA, length(numeric_cols)), 
                     eta_sq = rep(NA, length(numeric_cols)),
                     sem = rep(NA, length(numeric_cols)))

row.names(icc_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
  icc_df[numeric_cols[i], 'icc'] <- get_icc(numeric_cols[i])
  icc_df[numeric_cols[i], 'eta_sq'] <- get_eta(numeric_cols[i])
  icc_df[numeric_cols[i], 'sem'] <- get_sem(numeric_cols[i])
}

icc_df$dv = row.names(icc_df)
row.names(icc_df) = seq(1:nrow(icc_df))
icc_df$task = 'task'
icc_df[grep('survey', icc_df$dv), 'task'] = 'survey'
icc_df = icc_df %>%
  select(dv, task, icc, eta_sq, sem)
row.names(icc_df) = NULL

# Add column to tasks vs. surveys in correlation table
cor_df$dv = row.names(cor_df)
row.names(cor_df) = seq(1:nrow(cor_df))
cor_df$task = 'task'
cor_df[grep('survey', cor_df$dv), 'task'] = 'survey'
```

Average reliability metrics and the number of variables for both the surveys and the tasks
Both Spearman $\rho$'s as well the ICC's are higher for surveys.  
Systematic differences between the two measurements ($\eta^2$) are higher for tasks than surveys (potentially indicating learning effects?).  
Unexpectedly, median standard error of measurement is higher for surveys but the distribution below suggests that this is a side effect of having too many task measures with very low SEMs.
```{r}
datatable(rel_df %>%
  group_by(task) %>%
  summarise(median_spearman = median(spearman),
            median_icc = median(icc),
            median_eta_sq = median(eta_sq),
            median_sem = median(sem),
            num_vars = n()))
```

*Rank orders for tasks*

@@

*Rank order for surveys:*
- Demographics are excluded from this list of variables (because they are saved for predictive analyses).  
- Holt and Laury test-retest is the worst for *surveys*. But note that this is not a survey that has been studied extensively psychometrically like others. This is more similar to our tasks (one that is used more often by economists). In previous data I have worked with retest reliablities for this task was ~.35 so the results here are in line with this.  
- Otherwise the lowest survey reliabilities are 0.5.
```{r}

@@

  for(i in 1:nrow(ref_df)){
    for(j in 1:nrow(plot_df)){
      if(ref_df[i, bin_col] <= plot_df$xmax[j] & ref_df[i, bin_col]>=plot_df$xmin[j] & all(ref_df[i, group_col] == plot_df[j, group_col])){
        ref_df$bin[i] <- plot_df$bin[j]
      }
    }
  }

@@

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(cor_df %>%
                        ggplot(aes(spearman, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,cor_df, "spearman", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('Spearman')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

@@

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(icc_df %>%
                        ggplot(aes(icc, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,icc_df, "icc", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('ICC')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

@@

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(icc_df %>%
                        ggplot(aes(eta_sq, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,icc_df, "eta_sq", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('Partial Eta Squared')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

@@

```{r warning=FALSE, message=FALSE}
time_effect_vars = ggplot_build(icc_df %>%
               ggplot(aes(eta_sq, fill=task)) +
               geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
  mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
  get_labels(.,icc_df, "eta_sq", "task") %>%
  prettify_label(.) %>%
  filter(bin != 1 & bin !=2 & label != "NA") %>%
  select(label, task) %>%
  separate(label,into=c(as.character(1:27)),sep = "</br>") %>%
  gather(key, value, -task) %>%
  select(value, task) %>%
  na.omit()

@@

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(icc_df %>%
                        ggplot(aes(sem, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,icc_df, "sem", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('SEM')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

### Negative icc's?
```{r}
neg_icc_vars = icc_df %>%
  arrange(icc) %>% 
  filter(icc<0) %>%
  select(dv)
```

@@

```{r fig.height=12, fig.width=12}
high_icc_vars = icc_df %>%
  arrange(-icc)  %>% 
  filter(icc>0.92) %>%
  select(dv)

@@

```{r}
ggplotly(icc_df %>%
           left_join(cor_df[,c('spearman','dv')], by = 'dv') %>%
           ggplot(aes(spearman, icc, col=task, label=dv, label2=eta_sq))+
           geom_point()+
           theme_bw()+
           theme(legend.title = element_blank())+
           geom_abline(intercept = 0, slope=1))
```

@@

```{r}
datatable(icc_df %>%
  filter(icc>(-0.1), icc<0.1) %>%
  arrange(icc))
```

@@

```{r}
ggplotly(icc_df %>%
  ggplot(aes(eta_sq, icc, col=task, label=dv))+
  geom_point()+
  theme_bw()+
  theme(legend.title = element_blank()))
```

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -measure_type, -task, -ddm_task) %>%
  left_join(icc_df[,c("dv", "icc")], by='dv') %>%
  left_join(cor_df[,c("dv", "spearman")], by='dv') %>%
  separate(variable_type, c("label_1", "label_2", "label_3"), sep = ",") %>%
  mutate(label_1 = trim(label_1),
         label_2=trim(label_2),
         label_3=trim(label_3))%>%
  filter(label_1 %in% c('difference', 'EZ', 'hddm') ==FALSE & label_2 %in% c('difference', 'EZ', 'hddm') == FALSE) %>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.',remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  group_by(task_name) %>%
  summarise(mean_icc = mean(icc),
            mean_spearman = mean(spearman),
            num_measures = n(),
            num_trials = unique(num_all_trials)) %>%
  arrange(-mean_icc, -mean_spearman)

@@

```{r warning=FALSE, message=FALSE}
summary(lm(mean_icc ~ num_trials, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
summary(lm(mean_spearman ~ num_trials, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
summary(lm(mean_icc ~ num_measures, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
summary(lm(mean_spearman ~ num_measures, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
datatable(measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -variable_type, -task, -ddm_task) %>%
  left_join(icc_df[,c("dv", "icc")], by='dv') %>%
  left_join(cor_df[,c("dv", "spearman")], by='dv') %>%
  group_by(measure_type) %>%
  summarise(mean_icc = mean(icc),
            mean_spearman = mean(spearman),
            num_measures = n()) %>%
  arrange(-mean_icc, -mean_spearman))
```

@@

```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=10}
measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -variable_type, -task, -ddm_task) %>%
  left_join(icc_df[,c("dv", "icc")], by='dv') %>%
  left_join(cor_df[,c("dv", "spearman")], by='dv') %>%
  select(measure_type, icc, spearman) %>%
  gather(key, value, -measure_type) %>%
  ggplot(aes(value, fill=key))+
  geom_histogram(position='identity', alpha=0.5)+
  theme_bw()+
  facet_wrap(~measure_type)+
  theme(legend.title = element_blank())
```

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -measure_type, -task) %>%
  left_join(icc_df[,c("dv", "icc")], by='dv') %>%
  left_join(cor_df[,c("dv", "spearman")], by='dv') %>%
  separate(variable_type, c("label_1", "label_2", "label_3"), sep = ",") %>%
  mutate(label_1 = trim(label_1),
         label_2=trim(label_2),
         label_3=trim(label_3))%>%
  filter(ddm_task == 1) %>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.', remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  drop_na()
```

@@

```{r warning=FALSE, message=FALSE}
datatable(tmp %>%
  group_by(label_1,label_2, label_3) %>%
  summarise(mean_icc = mean(icc),
            mean_spearman = mean(spearman),
            n_vars = n()) %>%
  arrange(-mean_icc, -mean_spearman))
```

>>>>

```{r}
rel_df <- data.frame(spearman = rep(NA, length(numeric_cols)),
                     icc = rep(NA, length(numeric_cols)), 
                     eta_sq = rep(NA, length(numeric_cols)),
                     sem = rep(NA, length(numeric_cols)))

row.names(rel_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
  rel_df[numeric_cols[i], 'spearman'] <- get_spearman(numeric_cols[i]) 
  rel_df[numeric_cols[i], 'icc'] <- get_icc(numeric_cols[i])
  rel_df[numeric_cols[i], 'eta_sq'] <- get_eta(numeric_cols[i])
  rel_df[numeric_cols[i], 'sem'] <- get_sem(numeric_cols[i])
}

rel_df$dv = row.names(rel_df)
row.names(rel_df) = seq(1:nrow(rel_df))
rel_df$task = 'task'
rel_df[grep('survey', rel_df$dv), 'task'] = 'survey'
rel_df = rel_df %>%
  select(dv, task, spearman, icc, eta_sq, sem)
# row.names(rel_df) = NULL

Average reliability metrics and the number of variables for both the surveys and the tasks
Both Spearman $\rho$'s as well the ICC's are higher for surveys.  
Systematic differences between the two measurements ($\eta^2$) are higher for tasks than surveys (potentially indicating learning effects?).  
Unexpectedly, median standard error of measurement is higher for surveys but the distribution below suggests that this is a side effect of having too many task measures with very low SEMs.
```{r warning=FALSE, message=FALSE}
rel_df %>%
  group_by(task) %>%
  summarise(median_spearman = median(spearman),
            median_icc = median(icc),
            median_eta_sq = median(eta_sq),
            median_sem = median(sem),
            num_vars = n())
```

*Rank orders for reliabilities*

**ADD FILTER TO DATA TABLE FOR THIS DISPLAY INSTEAD OF MAKING TWO SEPARATE TABLES**

- Demographics are excluded from this list of variables (because they are saved for predictive analyses).  
- Holt and Laury test-retest is the worst for *surveys*. But note that this is not a survey that has been studied extensively psychometrically like others. This is more similar to our tasks (one that is used more often by economists). In previous data I have worked with retest reliablities for this task was ~.35 so the results here are in line with this.  
- Otherwise the lowest survey reliabilities are 0.5.

@@

@@

  for(i in 1:nrow(ref_df)){
    for(j in 1:nrow(plot_df)){
      # if(ref_df[i, bin_col] <= plot_df$xmax[j] & ref_df[i, bin_col]>=plot_df$xmin[j] & all(ref_df[i, group_col] == plot_df[j, group_col])){
      if(ref_df[i, bin_col] <= plot_df$xmax[j] & ref_df[i, bin_col]>=plot_df$xmin[j] & ref_df[i, group_col] == plot_df[j, group_col]){
        ref_df$bin[i] <- plot_df$bin[j]
      }
    }
  }

@@

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(rel_df %>%
                        ggplot(aes(spearman, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,rel_df, "spearman", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('Spearman')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

@@

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(rel_df %>%
                        ggplot(aes(icc, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,rel_df, "icc", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('ICC')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

@@

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(rel_df %>%
                        ggplot(aes(eta_sq, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,rel_df, "eta_sq", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('Partial Eta Squared')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

@@

```{r warning=FALSE, message=FALSE}
time_effect_vars = ggplot_build(rel_df %>%
               ggplot(aes(eta_sq, fill=task)) +
               geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
  mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
  get_labels(.,rel_df, "eta_sq", "task") %>%
  prettify_label(.) %>%
  filter(bin != 1 & bin !=2 & label != "NA") %>%
  select(label, task) %>%
  separate(label,into=c(as.character(1:27)),sep = "</br>") %>%
  gather(key, value, -task) %>%
  select(value, task) %>%
  na.omit()

@@

```{r warning=FALSE, message=FALSE}
ggplotly(ggplot_build(rel_df %>%
                        ggplot(aes(sem, fill=task)) +
                        geom_histogram(position = 'identity', alpha=0.5))$data[[1]] %>%
           mutate(task=ifelse(as.character(fill) == "#00BFC4", "task", "survey")) %>%
           get_labels(.,rel_df, "sem", "task") %>%
           prettify_label(.) %>%
           ggplot(aes(x,y, fill=task, label=label))+
           geom_bar(stat='identity', position='identity', alpha=0.5)+
           xlab('SEM')+
           ylab('Count')+
           theme_bw(),
         tooltip = 'label',
         hoverinfo = 'text')
```

### Negative icc's?
```{r}
neg_icc_vars = rel_df %>%
  arrange(icc) %>% 
  filter(icc<0) %>%
  select(dv)
```

@@

```{r fig.height=12, fig.width=12}
high_icc_vars = rel_df %>%
  arrange(-icc)  %>% 
  filter(icc>0.92) %>%
  select(dv)

@@

```{r}
ggplotly(rel_df %>%
           ggplot(aes(spearman, icc, col=task, label=dv, label2=eta_sq))+
           geom_point()+
           theme_bw()+
           theme(legend.title = element_blank())+
           geom_abline(intercept = 0, slope=1))
```

@@

```{r}
datatable(rel_df %>%
  filter(icc>(-0.1), icc<0.1) %>%
  arrange(icc))
```

@@

```{r}
ggplotly(rel_df %>%
  ggplot(aes(eta_sq, icc, col=task, label=dv))+
  geom_point()+
  theme_bw()+
  theme(legend.title = element_blank()))
```

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -measure_type, -task, -ddm_task) %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  separate(variable_type, c("label_1", "label_2", "label_3"), sep = ",") %>%
  mutate(label_1 = trim(label_1),
         label_2=trim(label_2),
         label_3=trim(label_3))%>%
  filter(label_1 %in% c('difference', 'EZ', 'hddm') ==FALSE & label_2 %in% c('difference', 'EZ', 'hddm') == FALSE) %>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.',remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  group_by(task_name) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            num_measures = n(),
            num_trials = unique(num_all_trials)) %>%
  arrange(-median_icc, -median_spearman)

@@

```{r warning=FALSE, message=FALSE}
summary(lm(median_icc ~ num_trials, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
summary(lm(median_spearman ~ num_trials, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
summary(lm(median_icc ~ num_measures, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
summary(lm(median_spearman ~ num_measures, data = tmp))
```

@@

```{r warning=FALSE, message=FALSE}
datatable(measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -variable_type, -task, -ddm_task) %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  group_by(measure_type) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            num_measures = n()) %>%
  arrange(-median_icc, -median_spearman))
```

@@

```{r warning=FALSE, message=FALSE, fig.width=12, fig.height=10}
measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -variable_type, -task, -ddm_task) %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  select(measure_type, icc, spearman) %>%
  gather(key, value, -measure_type) %>%
  ggplot(aes(value, fill=key))+
  geom_histogram(position='identity', alpha=0.5)+
  theme_bw()+
  facet_wrap(~measure_type)+
  theme(legend.title = element_blank())
```

@@

```{r warning=FALSE, message=FALSE}
tmp = measure_labels %>%
  mutate(dv = as.character(dv)) %>%
  filter(task == 'task') %>%
  select(-measure_description, -notes, -measure_type, -task) %>%
  left_join(rel_df[,c("dv", "spearman","icc")], by='dv') %>%
  separate(variable_type, c("label_1", "label_2", "label_3"), sep = ",") %>%
  mutate(label_1 = trim(label_1),
         label_2=trim(label_2),
         label_3=trim(label_3))%>%
  filter(ddm_task == 1) %>%
  separate(dv, c('task_name', 'extra_1', 'extra_2'), sep = '\\.', remove=FALSE) %>%
  select(-extra_1, -extra_2) %>%
  drop_na()
```

@@

```{r warning=FALSE, message=FALSE}
datatable(tmp %>%
  group_by(label_1,label_2, label_3) %>%
  summarise(median_icc = median(icc),
            median_spearman = median(spearman),
            n_vars = n()) %>%
  arrange(-median_icc, -median_spearman))
```
